---
title: Debt in Latin America
author: Joao RCB
date: '2021-04-26'
slug: debt-in-latin-america
categories:
  - R
  - debt
  - LatAm
tags: []
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=5.5, fig.height=3, fig.align = "center") 

```


Hello, people

Let's take a look at the Debt-to-GDP in Latin American using the imfr package.

The imfr package is a nice package that allow us to download data from the IMF's database inside R!

First we load our packages of interest.
```{r echo=TRUE, message=FALSE, warning=FALSE}
## IMF DATA
library(imfr) 

## for country codes. Very usefull!
library(countrycode) 

# Plots and data cleaning
library(lubridate) 
library(tidyverse)
library(hrbrthemes)
library(viridis)

```

Then we use imfr::imf_ids() to see the specific databases (Which is a lot).

I will use data from the HPDD (Historical Public Debt Data) and codes and names of countries from the contrycode package. 


First get the Codes:
```{r echo=TRUE, message=FALSE, warning=FALSE}
data(codelist)

country_set <- codelist
country_set<- country_set %>% 
  select(country.name.en , iso2c, iso3c, imf, continent, region) %>% filter(!is.na(imf) & !is.na(iso2c))


```



Then get the indicators of interest available at our database.
```{r echo=TRUE, message=FALSE, warning=FALSE}
imf_codes(codelist = "CL_INDICATOR_HPDD")

```

Getting the data
```{r echo=TRUE, message=FALSE, warning=FALSE}

HPDD <- tibble(
  imf_data(database_id = "HPDD",indicator ="GGXWDG_GDP",
           freq = 'A', start = 1950 , end = current_year())  ) 

HPDD <- HPDD %>% full_join(country_set, by = "iso2c") %>% 
  filter(region == "Latin America & Caribbean") 

HPDD %>% head(n=10)

```

Okokok... We have our data. Now lets sub select some countries of interest, do some cleaning and finally... plot!



```{r echo=TRUE, message=FALSE, warning=FALSE}

divida <- HPDD %>% 
  filter(iso3c %in% c("ARG","BOL", "BRA","CHL","COL","ECU","PER","PRY","URY","VEN")) %>% 
  mutate(date = year(
    as.Date(as.character(year), format = "%Y"  )) 
    )

tmp <- divida %>% ## this is for the plot
  mutate(iso3c2=iso3c)

```
  

Plot 1!

```{r echo=TRUE}

p1 <- divida%>% 
ggplot(aes(x=date,y= GGXWDG_GDP, group=iso3c, color=iso3c)) +
  geom_line( data=tmp %>% dplyr::select(-iso3c), aes(group=iso3c2), color="grey", size=2, alpha=0.5) +
  geom_line( aes(color=iso3c), color="black", size=2 )+
  theme(
    legend.position="none",
    panel.spacing = unit(0.1, "lines"),
    strip.text.x = element_text(size = 8),
    plot.title = element_text(size=14),
    panel.grid = element_blank()
  )  +
  ylab("") +
  ggtitle("Debt to GDP ration in LatAm") +
  theme_ipsum(grid = F) + 
  facet_wrap(iso3c~., ncol = 4)

print(p1)

```




Plot 2! 

```{r echo=TRUE}

p2 <- divida%>% 
  ggplot(aes(x=date,y= GGXWDG_GDP, group=iso3c, color=iso3c)) +
  geom_line(size=2) + 
  scale_colour_brewer(palette = "Set1") +
  ylab("") + xlab("") +
  ggtitle("Debt to GDP ration in LatAm") +
  theme_ipsum(grid = F) +
  theme(legend.title = element_blank(),
        legend.position = "bottom")


print(p2)
```



